---
title: "Poster analysis"
output: html_notebook
---

Set up environment
```{r}
#Set seed for random number generation
set.seed(444)

here::i_am("Poster_analysis.Rmd")
library(here)
```

Load in raw data for mice datasets
```{r}
m1k <- CreateObj(data.dir = here('PR1044_1000_mouse'), object.name = 'P1000_mouse_raw')
  
m2k <- CreateObj(data.dir = here('PR1044_2000_mouse'), object.name = 'P2000_mouse_raw')
  
m5k <- CreateObj(data.dir = here('PR0979_5000'), object.name = 'P5000_mouse_raw')

mouse.list <- list(m1k, m2k, m5k)
```

Mtio and QC for each dataset
```{r}
# Run Mito and QC
# Using 0.5 as cutoff
# NEEDS TO RUN IN CONSOLE

mouse.list_filtered <- lapply(mouse.list, MitoAndQC)

```

Normalise datasets and find variable features
```{r}
# Use SCTransform to normalise and find variable features
# Then run PCA and elbow plot
# Then run UMAP with user defined PCs

mouse.list_filtered <- lapply(mouse.list_filtered, SeuratBasics, transform = T)

```

Doublet removal

```{r}
# Remove doublets
# Plot cells removed
# Save plots and Seurat object
mouse.list_filtered <- lapply(mouse.list_filtered, DoubletRemoval, npcs = 50)

SaveObject(mouse.list_filtered, 'list_1k_2k_5k_afterDoubletRemoval')
```

Remove reductions and add orig.ident

```{r}
# Remove PCAs and UMAPs
mouse.list_filtered[[1]]@reductions$PCA <- NULL
mouse.list_filtered[[2]]@reductions$PCA <- NULL
mouse.list_filtered[[3]]@reductions$PCA <- NULL

mouse.list_filtered[[1]]@reductions$UMAP <- NULL
mouse.list_filtered[[2]]@reductions$UMAP <- NULL
mouse.list_filtered[[3]]@reductions$UMAP <- NULL

# Copy to new object
mouse.list_round2 <- mouse.list_filtered

# Add orig.ident for each dataset
mouse.list_round2[[1]]$orig.ident <- '1k'
mouse.list_round2[[2]]$orig.ident <- '2k'
mouse.list_round2[[3]]$orig.ident <- '5k'

```

Re-normalise and reduce dimensions on filtered dataset
```{r}
######################
# RUN IN CONSOLE
# As above but after QC and Doublet removal
mouse.list_round2 <- lapply(mouse.list_round2, SeuratBasics, transform = T)

```

Find integration anchors and integrate datasets

```{r}

# Select integration features and prep for integration
features <- SelectIntegrationFeatures(object.list = mouse.list_round2, nfeatures = 10000)

mouse.list_round2 <- PrepSCTIntegration(object.list = mouse.list_round2, anchor.features = features)

# Find integration anchors
anchors <- FindIntegrationAnchors(object.list = mouse.list_round2, normalization.method = "SCT",
    anchor.features = features)

# Integrate datasets
mouse <- IntegrateData(anchorset = anchors, normalization.method = "SCT")

# Set default assay
DefaultAssay(mouse) <- "integrated"

SaveObject(mouse, 'integrated_1k_2k_5k_mouse')
```

Create UMAP on integrated dataset

```{r}

#Seurat basics
mouse <- SeuratBasics(mouse)

SaveObject(mouse, 'integrated_1k_2k_5k_mouse_afterUMAP')

```

Plot UMAPs
```{r}

# Plot UMAP by origin
p <- DimPlot(mouse, reduction='umap', group.by = 'orig.ident') + NoAxes() + ggtitle('Mouse UMAP by dataset')

#Reorder factor to put ages in chronological order
mouse$sample <- factor(mouse$sample, levels = c('P0', 'P2', 'P4', 'P8', 'P11', 'P14', 'P21'))

# Plot UMAP by age
q <- DimPlot(mouse, reduction='umap', group.by = 'sample') + NoAxes() + ggtitle('Mouse UMAP by age')

# Plot UMAP by #reads
r <- FeaturePlot(mouse, reduction = 'umap', features = 'nCount_SCT') + NoAxes() + ggtitle('Mouse UMAP by nReads (normalised)')

# Plot UMAP by #features
s <- FeaturePlot(mouse, reduction = 'umap', features = 'nFeature_SCT') + NoAxes() + ggtitle('Mouse UMAP by nFeatures (normalised)')

#Plot all together
p + q + r + s

```

Clustering 

```{r}

mouse <- ReadObject('integrated_1k_2k_5k_mouse_afterUMAP')

mouse_clustered <- Clustering(mouse)

t <- DimPlot(mouse_clustered, reduction='umap', group.by = 'integrated_snn_res.0.2') + NoAxes() + ggtitle('Mouse UMAP by cluster')

SaveObject(mouse_clustered, 'integrated_1k_2k_5k_mouse_afterClustering')

p + q + r + s + t
```
Find all markers
```{r}

#u <- MarkerAnalysisPlot(mouse_clustered, final.res = final.res)

#Find markers
markers <-FindAllMarkers(mouse_clustered, 
                     test.use = 'MAST', 
                     group.by = 'integrated_snn_res.0.2')

SaveObject(markers, 'mouse_markers')
markers.list <- split(markers, markers$cluster)

top_markers <- lapply(markers.list, function(x) top_n(x, n=5, wt=avg_log2FC))
top_markers <- bind_rows(top_markers, .id = "id")

#Plot top 5 markers for mouse
top_plot <- sub("\\.\\d+$", "", rownames(top_markers))
top_plot <- unique(top_plot)
plot <- DotPlot(mouse_clustered, features = top_plot, group.by = resolution) + coord_flip()

plot

SaveFigure(plot, 'top5_mouse_markers', width = 20, height = 15)
```
Plot biological markers
```{r}
## SCs ##
# Otog - SCs & non-sensory epithelia
# Sparcl1 - type 2 SCs & vasuclar cells


## HCs ##
# Armc4 - all HCs
# Ush2a - all HCs, stronger in types 3-5
# Bdnf - HCs, opposite to Ush2a

## MES ##
# Pcca - mesenchyme
# Coch - mesenchyme
# Stim2 - non sensory epithelia

## GLIA ##
# Pllp
# Erc1 

## NEU ##
# Mbp - myelinated cells


biomarkers <- c("Myo7a",
                "Rasd2",
                "Otof",
 #               "Ocm", 
                "Plp1",
                "Lgr5",
 #               "Sox2",
  #              "Sparcl1",
                "Jag1"
#                "Tectb"
                )

plot <- FeaturePlot(mouse_clustered, features = biomarkers)
plot

SaveFigure(plot, "mouse_biomarkers_marcela", width = 8, height = 10)
```

Plot UMAP labelled by cell type
```{r}

#plt_clusters <- list(
#  c(2, 3, 4, 8, 10, 7),
#  c(0, 9),
#  c(1, 5),
#  c(6),
#  c(11, 13, 12)
#)

#plt_labels <- list(
#  "Supporting Cells",
#  "Mesenchyme",
#  "Hair Cells",
#  "Neurons & Glia",
#  "Non-sensory Epithelia"
#)

#Add labels and plot
mouse_labelled <- mouse_clustered

plt_labels <- c("Mesenchyme", "Hair Cells", rep("Supporting Cells", 3), "Hair Cells", "Neurons & Glia", 'Supporting Cells', 'Neurons & Glia', "Mesenchyme", 'Supporting Cells', rep('NS Epithelia', 3), rep('Neurons & Glia', 3))

names(plt_labels) <- levels(mouse_labelled)

mouse_labelled <- RenameIdents(mouse_labelled, plt_labels)

u <- DimPlot(mouse_labelled, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend() + NoAxes()

#SaveFigure(plot, 'mouse_labelled_umap_clusters', width = 8, height = 10)
#SaveObject(mouse_labelled, 'mouse_cluster_labels')
```

Plot all for poster
```{R}

# Plot UMAPs in order
p + q + t + r + s + u # Save as PDF 30 x 15 in

# Plot biomarkers
plot # Save as PDF 10 x 15 in

```

