---
title: "Poster analysis chick"
output: html_notebook
---

Set up environment
```{r}
#Set seed for random number generation
set.seed(444)

here::i_am("Poster_analysis_chick.Rmd")
library(here)
```

Load in raw data for mice datasets
```{r}
c1k <- CreateObj(data.dir = here('PR1044_1000_chick_gG5'), object.name = 'P1000_chick_raw')
  
c2k <- CreateObj(data.dir = here('PR1044_2000_chick_gG5'), object.name = 'P2000_chick_raw')
  
chick.list <- list(c1k, c2k)
```

Mtio and QC for each dataset
```{r}
# Run Mito and QC
# NEEDS TO RUN IN CONSOLE

chick.list_filtered <- lapply(chick.list, MitoAndQC)

```

Normalise datasets and find variable features
```{r}
# Use SCTransform to normalise and find variable features
# Then run PCA and elbow plot
# Then run UMAP with user defined PCs

chick.list_filtered <- lapply(chick.list_filtered, SeuratBasics, transform = T)

```

Doublet removal

```{r}
# Remove doublets
# Plot cells removed
# Save plots and Seurat object
chick.list_filtered <- lapply(chick.list_filtered, DoubletRemoval, npcs = 50)

SaveObject(chick.list_filtered, 'chick_list_1k_2k_afterDoubletRemoval')
```

Re-run normalisation on filtered dataset

```{r}
# Remove PCAs and UMAPs
chick.list_filtered[[1]]@reductions$PCA <- NULL
chick.list_filtered[[2]]@reductions$PCA <- NULL

chick.list_filtered[[1]]@reductions$UMAP <- NULL
chick.list_filtered[[2]]@reductions$UMAP <- NULL

# Copy to new object
chick.list_round2 <- chick.list_filtered

# Add orig.ident for each dataset
chick.list_round2[[1]]$orig.ident <- '1k'
chick.list_round2[[2]]$orig.ident <- '2k'

######################
# RUN IN CONSOLE
# As above but after QC and Doublet removal
chick.list_round2 <- lapply(chick.list_round2, SeuratBasics, transform = T)

```

Find integration anchors and integrate datasets

```{r}

# Select integration features and prep for integration
features <- SelectIntegrationFeatures(object.list = chick.list_round2, nfeatures = 10000)

chick.list_round2 <- PrepSCTIntegration(object.list = chick.list_round2, anchor.features = features)

# Find integration anchors
anchors <- FindIntegrationAnchors(object.list = chick.list_round2, normalization.method = "SCT",
    anchor.features = features)

# Integrate datasets
chick <- IntegrateData(anchorset = anchors, normalization.method = "SCT")

# Set default assay
DefaultAssay(chick) <- "integrated"

SaveObject(chick, 'integrated_1k_2k_chick')
```

Create UMAP on integrated dataset

```{r}

#Seurat basics
chick <- SeuratBasics(chick)

SaveObject(chick, 'integrated_1k_2k_chick_afterUMAP')

```

Plot UMAPs
```{r}
chick <- ReadObject('integrated_1k_2k_chick_afterUMAP')

# Plot UMAP by origin
p <- DimPlot(chick, reduction='umap', group.by = 'orig.ident') + NoAxes() + ggtitle('chick UMAP by dataset')

# Plot UMAP by age
q <- DimPlot(chick, reduction='umap', group.by = 'sample') + NoAxes() + ggtitle('chick UMAP by age')

# Plot UMAP by #reads
r <- FeaturePlot(chick, reduction = 'umap', features = 'nCount_SCT') + NoAxes() + ggtitle('chick UMAP by # of Reads')

# Plot UMAP by #features
s <- FeaturePlot(chick, reduction = 'umap', features = 'nFeature_SCT') + NoAxes() + ggtitle('chick UMAP by # of Features')

#Plot all together
p + q + r + s

SaveFigure(p + q + r + s, 'chick_integrated_umaps', width = 10, height = 8)
```

Clustering 

```{r}

#chick <- ReadObject('integrated_1k_2k_chick_afterUMAP')

chick_clustered <- Clustering(chick)

t <- DimPlot(chick_clustered, reduction='umap', group.by = 'integrated_snn_res.0.4') + NoAxes() + ggtitle('chick UMAP by cluster')

p + q + r + s + t

SaveObject(chick_clustered, 'integrated_1k_2k_chick_afterClustering')
SaveFigure(t, 'chick_umap_clustered', width = 10, height = 8)
```
Find all markers
```{r}

#u <- MarkerAnalysisPlot(chick_clustered, final.res = final.res)
final.res <- 0.4
resolution <- "integrated_snn_res.0.4"
#Find markers
chick.markers <-FindAllMarkers(chick_clustered, 
                     test.use = 'MAST', 
                     group.by = 'integrated_snn_res.0.4')

SaveObject(chick.markers, 'chick_markers')
chick.markers.list <- split(chick.markers, chick.markers$cluster)

chick.top_markers <- lapply(chick.markers.list, function(x) top_n(x, n=5, wt=avg_log2FC))
chick.top_markers <- bind_rows(chick.top_markers, .id = "id")

#Plot top 5 markers for chick
chick.top_plot <- sub("\\.\\d+$", "", rownames(chick.top_markers))
chick.top_plot <- unique(chick.top_plot)
plot <- DotPlot(chick_clustered, features = chick.top_plot, group.by = resolution) + coord_flip()

plot

SaveFigure(plot, 'top5_chick_markers', width = 20, height = 15)
```
Plot biological markers
```{r}
## SCs ##
# Otog - SCs & type 2 HCs
# Sparcl1 - type 2 HCs


## HCs ##
# Atoh1 - nascent hair cells
# Armc4 - all HCs
# Ush2a - all HCs
# Bdnf - all HCs

## MES ##
# Pcca - striolar supporting cells and all HCs
# Coch - extrastriolar & transitional supporting cells
# Stim2 - extrastriolar type 2 hair cells

## GLIA ##
# Pllp - extrastriolar & transitional supporting cells 
# Erc1 - striolar SCs, type 2 HCs

## NEU ##
# Mbp - striolar SCs
# SPOCK 1 - SCs and transitional SCs

chick_clustered <- ReadObject('integrated_1k_2k_chick_afterClustering')

chick.biomarkers <- str_to_upper(c("Myo7a",
#                "Rasd2",
                "Otof",
                "Ocm", 
#                "Plp1",
                "Lgr5",
                "Sox2",
#                "Sparcl1",
                "Jag1"
#                "Tectb"
                ))

plot <- FeaturePlot(chick_clustered, features = chick.biomarkers)
plot

SaveFigure(plot, "chick_biomarkers_marcela", width = 8, height = 10)

```

Plot UMAP labelled by cell type
```{r}

#Add labels and plot
chick_labelled <- chick_clustered

plt_labels <- c("Mesenchyme", "Supporting Cells", "Hair Cells", rep('Supporting Cells', 2), # 0 to 4
                'Neurons & Glia', 'Supporting Cells', 'Supporting Cells', 'Hair Cells', # 5 to 8
                'Mesenchyme', 'Neurons & Glia', 'Mesenchyme', 'Supporting Cells', # 9 to 12
                rep('Neurons & Glia', 2), 'Supporting Cells', 'Mesenchyme'
                )

names(plt_labels) <- levels(chick_labelled)

chick_labelled <- RenameIdents(chick_labelled, plt_labels)

u <- DimPlot(chick_labelled, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend() + NoAxes() + ggtitle('Chick UMAP by cell type')

SaveFigure(plot, 'chick_labelled_umap_clusters', width = 8, height = 10)
SaveObject(chick_labelled, 'chick_cluster_labels')
```


